Perturbation Theory in the Complex Plane: Exceptional Points and Where to Find Them

### 2021-01-28 Response_Letter/Response_Letter.tex

\documentclass[10pt]{letter}
\usepackage{UPS_letterhead,xcolor,mhchem,mathpazo,ragged2e,hyperref}
\newcommand{\alert}[1]{\textcolor{red}{#1}}
\definecolor{darkgreen}{HTML}{009900}

\begin{document}

\begin{letter}%
{To the Editors of the \textit{Journal of Physics: Condensed Matter}}

\opening{Dear Editors,}

\justifying
Please find attached a revised version of the manuscript entitled 
\begin{quote}
	\textit{``Perturbation Theory in the Complex Plane: Exceptional Points and Where to Find Them''}.
\end{quote}
We thank the reviewers for their constructive comments.
Our detailed responses to their comments can be found below.
For convenience, changes are highlighted in red in the revised version of the manuscript.

We look forward to hearing from you.

\closing{Sincerely, the authors.}

%%% REVIEWER 1 %%%
\noindent \textbf{\large Authors' answer to Reviewer \#1}

\begin{itemize}

	\item 
	{In this manuscript, the authors provide a comprehensive and thorough discussion of the role of exceptional point singularities in determining the performance of perturbation methods of electronic structure theory. The main subject  considered is the well-known Moller-Plesset (MP) theory to calculate the electronic correlation energy, and various resummation techniques aimed to improve the convergence of the MP series. The system used to illustrate the application of MP theory is the Hubbard dimer. The authors suggest how the combination of resummation methods and acceleration techniques, such as the Shanks transformation, can significantly improve the accuracy of pertubative approaches. The manuscript is well written and has the format of a very useful and informative review of the progress done in the field. For that reason, I was worried that it could not fit the audience of the Journal of Physics Condensed Matter journal, and it might be better placed in a different journal, focusing on reviews. Besides that point, I found some merit in the application of the Shanks transformation to different approximants for both the restricted and unrestricted versions of the theory, which focus on the corresponding set of orbitals. Thus, I am glad to recommend publication.}
	\\
	\alert{bla bla bla}

\end{itemize}

%%% REVIEWER 2 %%%
\noindent \textbf{\large Authors' answer to Reviewer \#2}

\begin{itemize}

	\item 
	{The topical review "Perturbation theory in the complex plane: Exceptional points and where to find them" reviews the origin and meaning of exceptional points in quantum chemistry. The Hubbard dimer is used throughout as an illustrative example for the restricted and unrestricted cases, exact solution (FCI), Hartree-Fock, and Rayleigh-Schrodinger perturbation theory. The authors review the ways that MP series in common quantum chemical applications can behave, using the classification schemes developed in the literature. Finally, they discuss alternative methods to approximate the Taylor series of RS theory by other functions which may have better, faster, or otherwise more robust convergence properties, including their own calculations based on the Shanks transformation.

This topical review is very well done and I recommend it for publication as is. I list my comments below for the authors to address if they choose. Because this manuscript is meant as a review and to have instructional value, my comments are a bit more pedagogical or curiousity driven than may be expected for a normal review. I am not an expert in this exact area, so my questions are meant to represent those a typical reader might have. My questions do not need to all be addressed before the manuscript is published.}
	\\
	\alert{bla bla bla}

\item 
{I
How am I to actually interpret the electronic energy E of such a non-Hermitian Hamiltonian? Is the energy associated with just the real part of the Riemann surface $E(\lambda)$ at every point, is it computable from some combination of Re and Im parts, or is the ordinary meaning of the energy lost completely? Is there a lifetime interpretation for the imaginary part of this non-Hermitian Hamiltonian in the wave function problem as in Green's function theory? Maybe something can be said about this because I suspect the imaginary part is somehow related to losses. This is hinted at in the introduction but was the first question that entered my mind.}
	\\
	\alert{bla bla bla}

\item 
{IIe
Could the authors make it very clear what is meant here by ``symmetry broken"? This term is seen and used very frequently but can be kind of ambiguous and rarely explained completely. What is the symmetry that is broken? Also, what exactly is meant by ``symmetry-pure" when referring to the molecular orbitals? This refers to equal composition of left and right sites, but maybe this can be said explicitly. A useful counterexample would be a mixing angle of zero, in which case the orbitals are either definitely left or definitely right.}
	\\
	\alert{bla bla bla}

\item 
{I personally can get confused because both spatial and spin symmetries are in play (or so I think). At high U/t, left and right sites are both open shell with one electron each. Am I to think of this as a broken spatial symmetry, broken spin symmetry, or both?}
	\\
	\alert{bla bla bla}

\item 
{It could be emphasized that Eqs. 20, 21a, and 21b are valid for both the RHF and UHF cases, which is why these equations lack an RHF or UHF subscript. Also, in the general case (either RHF or UHF), the bonding orbitals for the two sigma values are occupied and the two antibonding orbitals are unoccupied. Bonding and antibonding notation is not so common in the condensed matter literature and it may be missed that these correspond to occupied and empty single-particle states.}
	\\
	\alert{bla bla bla}

\item 
{IIf
What am I to think of values of $\lambda$ where $|\lambda|>1$? I guess this is mathematically well defined and useful for understanding $r_c$, but the physical system always corresponds to $|\lambda|=1$, even in the complex case. I suppose one needs to understand the singularities at other values of $\lambda$ in order to know the convergence of the series at the value we want, $\lambda = 1$. Some discussion could already be made that these values are not necessarily physical but determine the behavior of the physical series. Certain values away from $\lambda=1$ can also have physical meaning, like the autoionization process discussed later.}
	\\
	\alert{bla bla bla}

\item 
{IIIc
To be extra careful, should it be said that the RMP series is either convergent or divergent at the physical coupling value $|\lambda|=1$? I guess for either value of U, the series is convergent if you are close enough to the origin.}

The language describing the cylinder with unit radius could be improved from: "the radius of convergence indicated by the vertical cylinder of unit radius" I first thought the cylinder was marking the radius of convergence. It is meant to show the position of the EPs and indicate if the radius of convergence is $>1$ or $<1$. The language could be improved.
	\\
	\alert{bla bla bla}

IIId
\item 
{The classification of front-door and back-door intruder states is not totally clear. I raise this point because the journal is JPCM, and these terms are not common in condensed matter literature. Are intruder states equivalent to states creating exceptional points that ruin the series convergence?}
	\\
	\alert{bla bla bla}

\item 
{IIIe
Has "critical point" been defined yet?}
	\\
	\alert{bla bla bla}

\item 
{IIIf
The explanation of the critical point as the occurrence of the autoionization process is great and makes sense to me. I must say, though, I have trouble visualizing this in Figs. 7 and 8. I can trace any one curve continuously along the $\lambda$ axis, and even their ordering does not change. How does the critical point appear in these figures? Presumably, it is because the curves exactly intersect at the CP, but this does not on the surface appear so different than the avoided crossings related to the EPs discussed in earlier sections.

So what exactly in these figures indicates the suddent and abrupt change in the eigenstates that makes it a CP? Is it the derivative discontinuity in the energy on the real $\lambda$ axis, whereas the derivative on the real axis for an avoided crossing due to an exceptional point is continuous? Can this gradient discontinuity be used to define the CP? How do I see something like the autoionization in this figure? If this is correct, it could be emphasized a bit more by a comment referring specifically to that figure.

Is it possible to circle the CP as one does for an EP? Is there any meaning to this?}
	\\
	\alert{bla bla bla}

\item 
{IV
Can something more be said about how these methods are fitted (briefly, and in a general way)? For example, how does one fit the Pade coefficients? Presumably, the exact function is known to some order and/or inside of some radius of convergence. The approximant is fitted in this region to match the exact function and then only the approximant is used beyond the original radius of convergence or at higher order.}
	\\
	\alert{bla bla bla}

\item 
{Throughout the manuscript, the figures are excellent and really help the understanding. The authors have clearly taken time to prepare the manuscript and it can be published immediately.}
	\\
	\alert{bla bla bla}

\end{itemize}

\end{letter}
\end{document}

### 2021-01-29 Manuscript/EPAWTFT.tex
starting working on response letter

% Moller-Plesset
In electronic structure theory, the workhorse of time-independent perturbation theory is M\o{}ller--Plesset (MP) 
theory,\cite{Moller_1934} which remains one of the most popular methods for computing the electron 
correlation energy.\cite{Wigner_1934,Lowdin_1958}
This approach estimates the exact electronic energy by constructing a perturbative correction on top
of a mean-field Hartree--Fock (HF) approximation.\cite{SzaboBook}
The popularity of MP theory stems from its black-box nature, size-extensivity, and relatively low computational scaling, 
making it easily applied in a broad range of molecular research.\cite{HelgakerBook}
However, it is now widely recognised that the series of MP approximations (defined for a given perturbation
order $n$ as MP$n$) can show erratic, slow, or divergent behaviour that limit its systematic improvability.%
\cite{Laidig_1985,Knowles_1985,Handy_1985,Gill_1986,Laidig_1987,Nobes_1987,Gill_1988,Gill_1988a,Lepetit_1988} 
As a result, practical applications typically employ only the lowest-order MP2 approach, while 
the successive MP3, MP4, and MP5 (and higher order) terms are generally not considered to offer enough improvement
to justify their increased cost.
Turning the MP approximations into a convergent and 
systematically improvable series largely remains an open challenge.

@@

% MATHEMATICAL REPRESENTATION
Mathematically, Eq.~\eqref{eq:E_expansion} corresponds to a Taylor series expansion of the exact energy
around the reference system $\lambda = 0$.
The energy of the target ``physical'' system is recovered at the point $\lambda = 1$.
However, like all series expansions, Eq.~\eqref{eq:E_expansion} has a radius of convergence $\rc$. 
When $\rc < 1$, the Rayleigh--Schr\"{o}dinger expansion will diverge
for the physical system.
The value of $\rc$ can vary significantly between different systems and strongly depends on the particular decomposition
of the reference and perturbation Hamiltonians in Eq.~\eqref{eq:SchrEq-PT}.\cite{Mihalka_2017b}
%
% LAMBDA IN THE COMPLEX PLANE
From complex analysis, \cite{BenderBook} the radius of convergence for the energy can be obtained by looking for the 
non-analytic singularities of $E(\lambda)$ in the complex $\lambda$ plane.
This property arises from the following theorem: \cite{Goodson_2011}
\begin{quote}
\it
``The Taylor series about a point $z_0$ of a function over the complex $z$ plane will converge at a value $z_1$ 
if the function is non-singular at all values of $z$ in the circular region centred at $z_0$ with radius $\abs{z_1-z_0}$. 
If the function has a singular point $z_s$ such that $\abs{z_s-z_0} < \abs{z_1-z_0}$, 
then the series will diverge when evaluated at $z_1$.''
\end{quote}
As a result, the radius of convergence for a function is equal to the distance from the origin of the closest singularity
in the complex plane, referred to as the ``dominant'' singularity.
This singularity may represent a pole of the function, or a branch point (\eg, square-root or logarithmic)
in a multi-valued function.

@@

In the Hubbard dimer, the HF energy can be parametrised using two rotation angles $\ta$ and $\tb$ as
\begin{equation}
E_\text{HF}(\ta, \tb) = -t\, \qty( \sin \ta + \sin \tb ) + \frac{U}{2} \qty( 1 + \cos \ta \cos \tb ),
\end{equation}
where we have introduced bonding $\mathcal{B}^{\sigma}$ and antibonding $\mathcal{A}^{\sigma}$ molecular orbitals for 
the spin-$\sigma$ electrons as
\begin{subequations}
\begin{align}
    \mathcal{B}^{\sigma} & = \hphantom{-} \cos(\frac{\ts}{2}) \Lsi + \sin(\frac{\ts}{2}) \Rsi,
	\\
	\mathcal{A}^{\sigma} & = - \sin(\frac{\ts}{2}) \Lsi + \cos(\frac{\ts}{2}) \Rsi
\end{align}
\label{eq:RHF_orbs}
\end{subequations}
In the weak correlation regime $0 \le U \le 2t$, the angles which minimise the HF energy, 
\ie, $\pdv*{E_\text{HF}}{\ts} = 0$, are 
\begin{equation}
	\ta_\text{RHF} = \tb_\text{RHF} = \pi/2,
\end{equation}
giving the symmetry-pure molecular orbitals
\begin{align}
	\mathcal{B}_\text{RHF}^{\sigma} & = \frac{\Lsi + \Rsi}{\sqrt{2}},
	&
	\mathcal{A}_\text{RHF}^{\sigma} & = \frac{\Lsi - \Rsi}{\sqrt{2}},
\end{align}
and the ground-state RHF energy (Fig.~\ref{fig:HF_real})
\begin{equation}
	E_\text{RHF} \equiv E_\text{HF}(\ta_\text{RHF}, \tb_\text{RHF}) = -2t + \frac{U}{2}.
\end{equation}
However, in the strongly correlated regime $U>2t$, the closed-shell orbital restriction prevents RHF from 
modelling the correct physics with the two electrons on opposite sites.

@@

While most practical calculations generally consider only the MP2 or MP3 approximations, higher order terms can 
be computed to understand the convergence of the MP$n$ series.\cite{Handy_1985}
\textit{A priori}, there is no guarantee that this series will provide the smooth convergence that is desirable for a
systematically improvable theory.
In fact, when the reference HF wave function is a poor approximation to the exact wave function, 
for example in multi-configurational systems, MP theory can yield highly oscillatory, 
slowly convergent, or catastrophically divergent results.\cite{Gill_1986,Gill_1988,Handy_1985,Lepetit_1988,Leininger_2000}
Furthermore, the convergence properties of the MP series can depend strongly on the choice of restricted or
unrestricted reference orbitals.

Although practically convenient for electronic structure calculations, the MP partitioning is not 
the only possibility and alternative partitionings have been considered including: 
i) the Epstein-Nesbet (EN) partitioning which consists in taking the diagonal elements of $\hH$ as the zeroth-order Hamiltonian, \cite{Nesbet_1955,Epstein_1926} 
ii) the weak correlation partitioning in which the one-electron part is consider as the unperturbed Hamiltonian $\hH^{(0)}$ and the two-electron part is the perturbation operator $\hV$, and 
iii) the strong coupling partitioning where the two operators are inverted compared to the weak correlation partitioning. \cite{Seidl_2018,Daas_2020}
While an in-depth comparison of these different approaches can offer insight into 
their relative strengths and weaknesses for various situations, we will restrict our current discussion
to the convergence properties of the MP expansion.

@@

Olsen \etal\ investigated the causes of these divergences and the different types of convergence by
analysing the relation between the dominant singularity (\ie, the closest singularity to the origin) 
and the convergence behaviour of the series.\cite{Olsen_2000} 
Their analysis is based on Darboux's theorem: \cite{Goodson_2011}
\begin{quote}
\textit{``In the limit of large order, the series coefficients become equivalent to 
    the Taylor series coefficients of the singularity closest to the origin. ''}
\end{quote}
Following this theory, a singularity in the unit circle is designated as an intruder state, 
with a front-door (or back-door) intruder state if the real part of the singularity is positive (or negative).

@@

More recently, this two-state model has been extended to non-symmetric Hamiltonians as\cite{Olsen_2019}
\begin{equation}
	\underbrace{\mqty(\alpha & \delta_1 \\ \delta_2 & \beta)}_{\bH} = \underbrace{\mqty(\alpha & 0 \\ 0 & \beta + \gamma )}_{\bH^{(0)}} + \underbrace{\mqty( 0 & \delta_2 \\ \delta_1 & - \gamma)}_{\bV}.
\end{equation}
This extension allows various choices of perturbation to be analysed, including coupled cluster 
perturbation expansions \cite{Pawlowski_2019a,Pawlowski_2019b,Pawlowski_2019c,Pawlowski_2019d,Pawlowski_2019e} 
and other non-Hermitian perturbation methods.
Note that new forms of perturbation expansions only occur when the sign of $\delta_1$ and $\delta_2$ differ.
Using this non-Hermitian two-state model, the convergence of a perturbation series can be characterised 
according to a so-called ``archetype'' that defines the overall ``shape'' of the energy convergence.\cite{Olsen_2019} 
For Hermitian Hamiltonians, these archetypes can be subdivided into five classes 
(zigzag, interspersed zigzag, triadic, ripples, and geometric), 
while two additional archetypes (zigzag-geometric and convex-geometric) are observed in non-Hermitian Hamiltonians.
%
The geometric archetype appears to be the most common for MP expansions,\cite{Olsen_2019} but the 
ripples archetype corresponds to some of the early examples of MP convergence. \cite{Handy_1985,Lepetit_1988,Leininger_2000}
The three remaining Hermitian archetypes seem to be rarely observed in MP perturbation theory.
In contrast, the non-Hermitian coupled cluster perturbation theory,%
\cite{Pawlowski_2019a,Pawlowski_2019b,Pawlowski_2019c,Pawlowski_2019d,Pawlowski_2019e} exhibits a range of archetypes
including the interspersed zigzag, triadic, ripple, geometric, and zigzag-geometric forms.
This analysis highlights the importance of the primary singularity in controlling the high-order convergence, 
regardless of whether this point is inside or outside the complex unit circle. \cite{Handy_1985,Olsen_2000}

@@

Recently, Mih\'alka \etal\ have studied the effect of different partitionings, such as MP or EN theory, on the position of 
branch points and the convergence properties of Rayleigh--Schr\"odinger perturbation theory\cite{Mihalka_2017b} (see also
Ref.~\onlinecite{Surjan_2000}).
Taking the equilibrium and stretched water structures as an example, they estimated the radius of convergence using quadratic
Pad\'e approximants.
The EN partitioning provided worse convergence properties than the MP partitioning, which is believed to be
because the EN denominators are generally smaller than the MP denominators.
To remedy the situation, they showed that introducing a suitably chosen level shift parameter can turn a 
divergent series into a convergent one by increasing the magnitude of these denominators.\cite{Mihalka_2017b}
However, like the UMP series in stretched \ce{H2},\cite{Lepetit_1988} 
the cost of larger denominators is an overall slower rate of convergence.

@@

% HISTORICAL OVERVIEW
We began by presenting the fundamental concepts behind non-Hermitian extensions of quantum chemistry into the complex plane, 
including the Hartree--Fock approximation and Rayleigh--Schr\"odinger perturbation theory.
We then provided a comprehensive review of the various research that has been performed 
around the physics of complex singularities in perturbation theory, with a particular focus on M{\o}ller--Plesset theory. 
Seminal contributions from various research groups have revealed highly oscillatory,
slowly convergent, or catastrophically divergent behaviour of the restricted and/or unrestricted MP perturbation series.%
\cite{Laidig_1985,Knowles_1985,Handy_1985,Gill_1986,Laidig_1987,Nobes_1987,Gill_1988,Gill_1988a,Lepetit_1988} 
In particular, the spin-symmetry-broken unrestricted MP series is notorious
for giving incredibly slow convergence.\cite{Gill_1986,Nobes_1987,Gill_1988a,Gill_1988}
All these behaviours can be rationalised and explained by the position of exceptional points
and other singularities that arise when perturbation theory is extended across the complex plane.

>>>>

% Moller-Plesset
In electronic structure theory, the workhorse of time-independent perturbation theory is M\o{}ller--Plesset (MP) 
theory,\cite{Moller_1934} which remains one of the most popular methods for computing the electron 
correlation energy.\cite{Wigner_1934,Lowdin_1958}
This approach estimates the exact electronic energy by constructing a perturbative correction on top
of a mean-field Hartree--Fock (HF) approximation.\cite{SzaboBook}
The popularity of MP theory stems from its black-box nature, size-extensivity, and relatively low computational scaling, 
making it easily applied in a broad range of molecular research.\cite{HelgakerBook}
However, it is now widely recognised that the series of MP approximations (defined for a given perturbation
order $n$ as MP$n$) can show erratic, slow, or divergent behaviour that limit its systematic improvability.%
\cite{Laidig_1985,Knowles_1985,Handy_1985,Gill_1986,Laidig_1987,Nobes_1987,Gill_1988,Gill_1988a,Lepetit_1988,Malrieu_2003} 
As a result, practical applications typically employ only the lowest-order MP2 approach, while 
the successive MP3, MP4, and MP5 (and higher order) terms are generally not considered to offer enough improvement
to justify their increased cost.
Turning the MP approximations into a convergent and 
systematically improvable series largely remains an open challenge.

@@

% MATHEMATICAL REPRESENTATION
Mathematically, Eq.~\eqref{eq:E_expansion} corresponds to a Taylor series expansion of the exact energy
around the reference system $\lambda = 0$.
The energy of the target ``physical'' system is recovered at the point $\lambda = 1$.
However, like all series expansions, Eq.~\eqref{eq:E_expansion} has a radius of convergence $\rc$. 
When $\rc < 1$, the Rayleigh--Schr\"{o}dinger expansion will diverge
for the physical system.
The value of $\rc$ can vary significantly between different systems and strongly depends on the particular decomposition
of the reference and perturbation Hamiltonians in Eq.~\eqref{eq:SchrEq-PT}.\cite{Mihalka_2017b}
%
% LAMBDA IN THE COMPLEX PLANE
From complex analysis, \cite{BenderBook} the radius of convergence for the energy can be obtained by looking for the 
non-analytic singularities of $E(\lambda)$ in the complex $\lambda$ plane.
This property arises from the following theorem: \cite{Goodson_2011}
\begin{quote}
\it
``The Taylor series about a point $z_0$ of a function over the complex $z$ plane will converge at a value $z_1$ 
if the function is non-singular at all values of $z$ in the circular region centred at $z_0$ with radius $\abs{z_1-z_0}$. 
If the function has a singular point $z_s$ such that $\abs{z_s-z_0} < \abs{z_1-z_0}$, 
then the series will diverge when evaluated at $z_1$.''
\end{quote}
As a result, the radius of convergence for a function is equal to the distance from the origin of the closest singularity
in the complex plane, referred to as the ``dominant'' singularity.
This singularity may represent a pole of the function, or a branch point (\eg, square-root or logarithmic)
in a multi-valued function.
\titou{T2: define here critical point.}

@@

In the Hubbard dimer, the HF energy can be parametrised using two rotation angles $\ta$ and $\tb$ as
\begin{equation}
\label{eq:EHF}
E_\text{HF}(\ta, \tb) = -t\, \qty( \sin \ta + \sin \tb ) + \frac{U}{2} \qty( 1 + \cos \ta \cos \tb ),
\end{equation}
where we have introduced \titou{occupied $\psi_1^{\sigma}$} and \titou{unoccupied $\psi_2^{\sigma}$} molecular orbitals for 
the spin-$\sigma$ electrons as
\begin{subequations}
\begin{align}
	\label{eq:psi1}
    \titou{\psi_1^{\sigma}} & = \hphantom{-} \cos(\frac{\ts}{2}) \Lsi + \sin(\frac{\ts}{2}) \Rsi,
	\\
	\label{eq:psi2}
	\titou{\psi_2^{\sigma}} & = - \sin(\frac{\ts}{2}) \Lsi + \cos(\frac{\ts}{2}) \Rsi
\end{align}
\label{eq:RHF_orbs}
\end{subequations}
\titou{Equations \eqref{eq:EHF}, \eqref{eq:psi1}, and \eqref{eq:psi2} are valid for both RHF and UHF.}
In the weak correlation regime $0 \le U \le 2t$, the angles which minimise the HF energy, 
\ie, $\pdv*{E_\text{HF}}{\ts} = 0$, are 
\begin{equation}
	\ta_\text{RHF} = \tb_\text{RHF} = \pi/2,
\end{equation}
giving the symmetry-pure molecular orbitals
\begin{align}
	\titou{\psi_{1,\text{RHF}}^{\sigma}} & = \frac{\Lsi + \Rsi}{\sqrt{2}},
	&
	\titou{\psi_{2,\text{RHF}}^{\sigma}} & = \frac{\Lsi - \Rsi}{\sqrt{2}},
\end{align}
and the ground-state RHF energy (Fig.~\ref{fig:HF_real})
\begin{equation}
	E_\text{RHF} \equiv E_\text{HF}(\ta_\text{RHF}, \tb_\text{RHF}) = -2t + \frac{U}{2}.
\end{equation}
However, in the strongly correlated regime $U>2t$, the closed-shell orbital restriction prevents RHF from 
modelling the correct physics with the two electrons on opposite sites.

@@

While most practical calculations generally consider only the MP2 or MP3 approximations, higher order terms can 
be computed to understand the convergence of the MP$n$ series.\cite{Handy_1985}
\textit{A priori}, there is no guarantee that this series will provide the smooth convergence that is desirable for a
systematically improvable theory.
In fact, when the reference HF wave function is a poor approximation to the exact wave function, 
for example in multi-configurational systems, MP theory can yield highly oscillatory, 
slowly convergent, or catastrophically divergent results.\cite{Gill_1986,Gill_1988,Handy_1985,Lepetit_1988,Leininger_2000,Malrieu_2003}
Furthermore, the convergence properties of the MP series can depend strongly on the choice of restricted or
unrestricted reference orbitals.

Although practically convenient for electronic structure calculations, the MP partitioning is not 
the only possibility and alternative partitionings have been considered \cite{Surjan_2004} including: 
i) the Epstein-Nesbet (EN) partitioning which consists in taking the diagonal elements of $\hH$ as the zeroth-order Hamiltonian, \cite{Nesbet_1955,Epstein_1926} 
ii) the weak correlation partitioning in which the one-electron part is consider as the unperturbed Hamiltonian $\hH^{(0)}$ and the two-electron part is the perturbation operator $\hV$, and 
iii) the strong coupling partitioning where the two operators are inverted compared to the weak correlation partitioning. \cite{Seidl_2018,Daas_2020}
While an in-depth comparison of these different approaches can offer insight into 
their relative strengths and weaknesses for various situations, we will restrict our current discussion
to the convergence properties of the MP expansion.

@@

Olsen \etal\ investigated the causes of these divergences and the different types of convergence by
analysing the relation between the dominant singularity (\ie, the closest singularity to the origin) 
and the convergence behaviour of the series.\cite{Olsen_2000} 
Their analysis is based on Darboux's theorem: \cite{Goodson_2011}
\begin{quote}
\textit{``In the limit of large order, the series coefficients become equivalent to 
    the Taylor series coefficients of the singularity closest to the origin. ''}
\end{quote}
\titou{Following this theory, a singularity in the unit circle is designated as an intruder state, 
with a front-door (or back-door) intruder state if the real part of the singularity is positive (or negative).}

@@

More recently, this two-state model has been extended to non-symmetric Hamiltonians as\cite{Olsen_2019}
\begin{equation}
	\underbrace{\mqty(\alpha & \delta_1 \\ \delta_2 & \beta)}_{\bH} = \underbrace{\mqty(\alpha & 0 \\ 0 & \beta + \gamma )}_{\bH^{(0)}} + \underbrace{\mqty( 0 & \delta_2 \\ \delta_1 & - \gamma)}_{\bV}.
\end{equation}
This extension allows various choices of perturbation to be analysed, including coupled cluster 
perturbation expansions \cite{Pawlowski_2019a,Pawlowski_2019b,Pawlowski_2019c,Pawlowski_2019d,Pawlowski_2019e} 
and other non-Hermitian perturbation methods.
Note that new forms of perturbation expansions only occur when the sign of $\delta_1$ and $\delta_2$ differ.
Using this non-Hermitian two-state model, the convergence of a perturbation series can be characterised 
according to a so-called ``archetype'' that defines the overall ``shape'' of the energy convergence.\cite{Olsen_2019} 
For Hermitian Hamiltonians, these archetypes can be subdivided into five classes 
(zigzag, interspersed zigzag, triadic, ripples, and geometric), 
while two additional archetypes (zigzag-geometric and convex-geometric) are observed in non-Hermitian Hamiltonians.
%
The geometric archetype appears to be the most common for MP expansions,\cite{Olsen_2019} but the 
ripples archetype corresponds to some of the early examples of MP convergence. \cite{Handy_1985,Lepetit_1988,Leininger_2000,Malrieu_2003}
The three remaining Hermitian archetypes seem to be rarely observed in MP perturbation theory.
In contrast, the non-Hermitian coupled cluster perturbation theory,%
\cite{Pawlowski_2019a,Pawlowski_2019b,Pawlowski_2019c,Pawlowski_2019d,Pawlowski_2019e} exhibits a range of archetypes
including the interspersed zigzag, triadic, ripple, geometric, and zigzag-geometric forms.
This analysis highlights the importance of the primary singularity in controlling the high-order convergence, 
regardless of whether this point is inside or outside the complex unit circle. \cite{Handy_1985,Olsen_2000}

@@

Recently, Mih\'alka \etal\ have studied the effect of different partitionings, such as MP or EN theory, on the position of 
branch points and the convergence properties of Rayleigh--Schr\"odinger perturbation theory\cite{Mihalka_2017b} (see also
Refs.~\onlinecite{Szabados_1999,Surjan_2000,Szabados_2003}).
Taking the equilibrium and stretched water structures as an example, they estimated the radius of convergence using quadratic
Pad\'e approximants.
The EN partitioning provided worse convergence properties than the MP partitioning, which is believed to be
because the EN denominators are generally smaller than the MP denominators.
To remedy the situation, they showed that introducing a suitably chosen level shift parameter can turn a 
divergent series into a convergent one by increasing the magnitude of these denominators.\cite{Mihalka_2017b}
However, like the UMP series in stretched \ce{H2},\cite{Lepetit_1988} 
the cost of larger denominators is an overall slower rate of convergence.

@@

% HISTORICAL OVERVIEW
We began by presenting the fundamental concepts behind non-Hermitian extensions of quantum chemistry into the complex plane, 
including the Hartree--Fock approximation and Rayleigh--Schr\"odinger perturbation theory.
We then provided a comprehensive review of the various research that has been performed 
around the physics of complex singularities in perturbation theory, with a particular focus on M{\o}ller--Plesset theory. 
Seminal contributions from various research groups have revealed highly oscillatory,
slowly convergent, or catastrophically divergent behaviour of the restricted and/or unrestricted MP perturbation series.%
\cite{Laidig_1985,Knowles_1985,Handy_1985,Gill_1986,Laidig_1987,Nobes_1987,Gill_1988,Gill_1988a,Lepetit_1988,Malrieu_2003} 
In particular, the spin-symmetry-broken unrestricted MP series is notorious
for giving incredibly slow convergence.\cite{Gill_1986,Nobes_1987,Gill_1988a,Gill_1988}
All these behaviours can be rationalised and explained by the position of exceptional points
and other singularities that arise when perturbation theory is extended across the complex plane.
